Group Signatures Meet Anonymous Credentials: Or, How to Build Privacy Preserving Identities with Accountability

### 2022-02-09 prism-backend/research/gsac/introduction.tex
First complete version of the model.
Some minor edits.

In this work, we combine two closely related schemes -- group signatures (GS)
and anonymous credentials (AC) -- which, for some reason, have followed a
separate path up to now, beyond individual constructions from both fields
drawing inspiration (and techniques) from one another. This combination is in
some sense a natural evolution, which we do believe may have relevant impact in
practical settings. On the one hand, anonymous credentials enable a richer
functionality in regards to what one can prove about his own identity, in a
privacy-preserving way. Academic proof of this is the fact that AC systems have
naturally evolved to enabling arbitrary claims that only reveal the minimally
possible 1-bit of information \needcite. Also, AC systems have been proposed to
go even beyond what one can prove about his identity, covering also orthogonal
functionality such as delegation \needcite, or aggregation of multiple
credentials into one \needcite. Even outside academia, prominent
technical workgroups \needcite, startups \needcite, and big techs \needcite are
promoting and adopting solutions which are directly inspired by AC systems. On
the other hand, GS 
schemes have inspection functionality (e.g., opening and linking of signatures)
at their core, which is reflected formally through non-frameability properties.
This has originated a large body of research focusing on the construction and
formalisation of variants of these notions \needcite. Consequently, group
signatures tend to be proposed for scenarios that require some sort of
de-anonymization \needcite. Indeed, this is not possible in ACs, which at most
support revocation of issued credentials, which essentially can compare to
linking, but not to opening-related notions.

Our goal is to get the best of both worlds: leverage the richer functionality
of ACs in what respects to proving, in a privacy-preserving manner, useful
claims on one's identity; and, at the same time support, through inspection
capabilities, use cases that may require some sort of de-anonymiztion. We
could target the ideal case of allowing to create proofs of arbitrary predicates
on one's identity (as some AC systems do), as well as provide flexible or several
notions of non-frameability. However, this being -- as far as we know -- the
first work that explores this intersection, we opt for keeping things simple
for the sake of facilitating technical precission and readability, and fostering
discussion. Therefore, we focus on the case of selective attribute disclosure
for ACs, and conventional opening for GSs, building a GSAC scheme that supports
both.

>>>>

In this work, we combine two closely related schemes -- group signatures (GS)
and anonymous credentials (AC) -- which have followed a
separate path up to now, beyond individual constructions from both fields
drawing inspiration (and techniques) from one another. This combination is in
some sense a natural evolution, which we do believe may have relevant impact in
practical settings. On the one hand, anonymous credentials enable a richer
functionality in regards to what one can prove about his own identity, in a
privacy-preserving way. Academic proof of this is the fact that AC systems have
naturally evolved to enabling arbitrary claims that only reveal the minimally
possible 1-bit of information \needcite. Also, AC systems have been proposed to
go even beyond what one can prove about his identity, covering also orthogonal
functionality such as delegation \needcite, or aggregation of multiple
credentials into one \needcite. Even outside academia, prominent
technical workgroups \needcite, startups \needcite, and big techs \needcite are
promoting and adopting solutions which are directly inspired by AC systems. On
the other hand, GS 
schemes have inspection functionality (e.g., opening and linking of signatures)
at their core, which is reflected formally through non-frameability properties.
This has originated a large body of research focusing on the construction and
formalisation of variants of these notions \needcite. Consequently, group
signatures tend to be proposed for scenarios that require some sort of
de-anonymization \needcite. Indeed, this is not possible in ACs, which at most
support revocation of issued credentials, which essentially can compare to
linking, but not to opening-related notions.

Our goal is to get the best of both worlds: leverage the richer functionality
of ACs in what respects to proving, in a privacy-preserving manner, useful
claims on one's identity; and, at the same time support, through inspection
capabilities, use cases that may require some sort of de-anonymiztion. We
could target the ideal case of allowing to create proofs of arbitrary predicates
on one's identity (as some AC systems do), as well as provide flexible or several
notions of non-frameability. However, this being -- as far as we know -- the
first work that explores this intersection, we opt for keeping things simple
for the sake of facilitating technical precission and readability, and fostering
discussion. Therefore, we focus on the case of selective attribute disclosure
for ACs, and conventional opening for GSs, building a GSAC scheme that supports
both. \commentwho{Jesus}{This being the first work in this direction is not (at
  least in part) true. See the related work section. Then update this
  accordingly.}

### 2022-02-09 prism-backend/research/gsac/model.tex
Some minor edits.

The first choice that has to be made is whether to make sign/show and
verification a non-interactive or an interactive process -- in group signatures,
we are in the former case, whereas in anonymous credentials, we are in the
latter. While both options are of course valid, this has direct impact in the
modelling. We choose the non-interactive approach, and then prove how to
generically translate it into an interactive process. The reason is that this
results in a more flexible building block, suitable for a wider set of
scenarios. And, syntactically, it seems more natural to combine two
non-interactive processes into an interactive one, than the other way around.

@@

Getting closer to the anonymity definition of anonymous credentials also has the
side effect to introduce the notions of attributes for free into the domain of
group signatures -- which, up to now, had considered membership credentials
\emph{without} attributes. Yet another side effect of introducing multiple
attributes (if we think from a group signature perspective) is that, in order
to avoid trivial wins, the anonymity game has to enforce that both challenge
credentials reveal the same predicate on their attributes.

@@

\paragraph{Global Variables.} %
All the honestly generated users (i.e., all honestly generated user key pairs,
since we assume a one-to-one relationship between user and user key pair), as
well as all honestly generated credentials, are assigned identifiers. That we
typically write as \uid for users, and \cid for credentials. The adversary can
refer to any individual user or credential using the corresponding identifier --
even though he may not know the actual contents of the key pairs or credentials.
In games that involve challenge user or credential identifiers, we use \cuid and
\ccid to refer to these challenge user/credential.
%
The games, usually through the oracles, keep track of the users (through
set \USK) and credentials (through table \CRED) that are created as a result of
oracle calls by the adversary, as well as the attributes that each credential
is associated (through table \ATTR), and the user to which each credential has
been issued to (through table \OWNR). We use the user and credential identifiers
to reference specific user key pairs, credentials, and credential attributes.
For instance, $\USK[\uid]$ refers to the user key pair corresponding to user
with identifier \uid, $\PUBUSK[\uid]$ refers to the public key of that key pair,
and $\PRVUSK[\uid]$ to the private key; credential $\CRED[\cid]$ refers to the
credential data associated to the credential with identifier \cid; $\ATTR[\cid]$
is the set of attributes that was assigned to the credential with identifier
\cid; and $\OWNR[\cid]$ is set to the \uid corresponding to the user that
credential \cid was issued to.
%
Additionally, the games keep track of all honest and corrupt users that have
been generated, through sets \HU and \CU, respectively. They also keep track of
the signatures that have been honestly produced, through the \SIG table. Since
signatures are produced through credentials, the \SIG table is indexed with
{\cid}s. For the anonymity game, we also need to keep track of the challenge
signatures that the adversary has obtained, in order to prevent trivial wins
by allowing any of them to be opened. 
%
All global variables are initially set by the games to undefined values (denoted
with $\bot$, and all tables/sets are initialized as empty tables/sets, denoted
with $\empty$. Also, for readability, we abuse the syntax as follows: we write
$\CRED[\uid]$ to mean $\CRED[\cid]$ for all $\cid$ such that
$\OWNR[\cid] = \uid$; \todo{Something else?}

@@

>>>>

The first choice that has to be made is whether to make sign/show and
verification a non-interactive or an interactive process -- in group signatures,
we are in the former case, whereas in anonymous credentials, we are in the
latter. While both options are of course valid, this has direct impact in the
modelling. We choose the non-interactive approach, and then prove how to
generically translate it into an interactive process \todo{do we?}. The reason
is that this results in a more flexible building block, suitable for a wider set
of scenarios. And, syntactically, it seems more natural to combine two
non-interactive processes into an interactive one, than the other way around.

@@

Getting closer to the anonymity definition of anonymous credentials also has the
side effect to introduce the notions of attributes for free into the domain of
group signatures -- which, up to now, had considered membership credentials
\emph{without} attributes. Yet another side effect of introducing multiple
attributes (if we think from a group signature perspective) is that, in order
to avoid trivial wins, the anonymity game has to enforce that both challenge
credentials reveal the same predicate on their attributes.
\commentwho{Jesus}{Again, taking into account ABGS, this is no longer true.
  Cross-check and re-write.}

@@

\paragraph{Global Variables.} %
All the honestly generated users (i.e., all honestly generated user key pairs,
since we assume a one-to-one relationship between user and user key pair), as
well as all honestly generated credentials, are assigned identifiers. Wex
typically write as \uid for users' identifiers, and \cid for credentials'. The
adversary can
refer to any individual user or credential using the corresponding identifier --
even though he may not know the actual contents of the key pairs or credentials.
In games that involve challenge user or credential identifiers, we use \cuid and
\ccid to refer to these challenge user/credential.
%
The games, usually through the oracles, keep track of the users (through
set \USK) and credentials (through table \CRED) that are created as a result of
oracle calls by the adversary, as well as the attributes that each credential
is associated (through table \ATTR), and the user to which each credential has
been issued to (through table \OWNR). We use the user and credential identifiers
to reference specific user key pairs, credentials, and credential attributes.
For instance, $\USK[\uid]$ refers to the user key pair corresponding to user
with identifier \uid, $\PUBUSK[\uid]$ refers to the public key of that key pair,
and $\PRVUSK[\uid]$ to the private key; credential $\CRED[\cid]$ refers to the
credential data associated to the credential with identifier \cid; $\ATTR[\cid]$
is the set of attributes that was assigned to the credential with identifier
\cid; and $\OWNR[\cid]$ is set to the \uid corresponding to the user that
credential \cid was issued to.
%
Additionally, the games keep track of all honest and corrupt users that have
been generated, through sets \HU and \CU, respectively. They also keep track of
the signatures that have been honestly produced, through the \SIG table. Since
signatures are produced through credentials, the \SIG table is indexed with
{\cid}s. For the anonymity game, we also need to keep track of the challenge
signatures that the adversary has obtained, in order to prevent trivial wins
by allowing any of them to be opened. 
%
All global variables are initially set by the games to empty values (denoted
with $\bot$), and all tables/sets are initialized as empty tables/sets, denoted
with $\emptyset$. Also, for readability, we abuse the syntax as follows: we
write $\CRED[\uid]$ to mean $\CRED[\cid]$ for all $\cid$ such that
$\OWNR[\cid] = \uid$; \todo{Something else?}

@@

\todo{It would be nice to prove that a \GSAC scheme to which we remove the
  \Open/\Judge functions becomes an AC scheme. And conversely, a \GSAC scheme
  where all credentials have no attributes, and where we restrict to only
  one credential per user, becomes a conventional GS scheme.}
